A causal analysis of the effect of abortion policies on maternal mortality

### 2022-06-08 PaperAnalysis.Rmd
date_update

---
title: "A causal analysis of the effect of abortion policies on maternal mortality"
author: "Helena Bakic, Anna-Lisa Wirth, Ángela Duarte Pardo"
date: "5/23/2022"
output: html_document
bibliography: references.bib
---

@@

Abortion law and the negative consequences of the lack of safe abortions are highly contentious policy issues that have been making headlines this year. Further, the [Sustainable Development Goal indicator 3.1](https://www.who.int/data/gho/indicator-metadata-registry/imr-details/26) of the United Nations to reduce Maternal Death to less than 70 per 100,000 live births by 2030 relies on access to safe and legal abortions. In the paper we are analysing, the authors are testing whether and, if so, to what extent abortion law as a proxy for access to safe abortions causally affects maternal mortality rates in over 160 countries around the globe. Establishing a causal relation could strengthen the implementation of abortion liberalisation policy. This is especially needed in a time where several states already withdrew or are withdrawing access to safe abortions on many grounds, such as [Poland](https://www.amnesty.org/en/latest/news/2021/10/poland-a-year-on-abortion-ruling-harms-women/) and the [USA](https://www.euronews.com/2022/05/03/us-supreme-court-moving-to-overturn-key-abortion-rights-law-says-leaked-report). Conversly, some other countries, for instance, [France](https://www.hrw.org/news/2022/03/01/france-expands-abortion-access-two-key-moves) are planning to expand abortion rights in 2022. So does abortion law flexibility cause maternal mortality? We will be taking a close critical look at a paper that tries to establish exactly that and also run our own data analysis to address this question.

>>>>

---
title: "A causal analysis of the effect of abortion policies on maternal mortality"
author: "Helena Bakic, Anna-Lisa Wirth, Ángela Duarte Pardo"
date: "6/08/2022"
output: html_document
bibliography: references.bib
---

@@

Abortion law and the negative consequences of the lack of safe abortions are highly contentious policy issues that have been making headlines this year. Further, the [Sustainable Development Goal indicator 3.1](https://www.who.int/data/gho/indicator-metadata-registry/imr-details/26) of the United Nations to reduce Maternal Death to less than 70 per 100,000 live births by 2030 relies on access to safe and legal abortions. In the paper we are analysing, the authors are testing whether and, if so, to what extent abortion law as a proxy for access to safe abortions causally affects maternal mortality rates in over 160 countries around the globe. Establishing a causal relation could strengthen the implementation of abortion liberalisation policy. This is especially needed in a time where several states already withdrew or are withdrawing access to safe abortions on many grounds, such as [Poland](https://www.amnesty.org/en/latest/news/2021/10/poland-a-year-on-abortion-ruling-harms-women/) and the [USA](https://www.euronews.com/2022/05/03/us-supreme-court-moving-to-overturn-key-abortion-rights-law-says-leaked-report). Conversely, some other countries, for instance, [France](https://www.hrw.org/news/2022/03/01/france-expands-abortion-access-two-key-moves) are planning to expand abortion rights in 2022. So does abortion law flexibility cause maternal mortality? We will be taking a close critical look at a paper that tries to establish exactly that and also run our own data analysis to address this question.
